Recommender System Based on Matrix Factorization

### 2019-11-18 simple_ml/example/report5/paper.tex
complete report5..

% Define document title and author
	\title{Adaboost for face image classification}
	\maketitle

% Write abstract here
\begin{abstract}
In this report, we solve the face/non-face classification problem by adaboost with Decision Tree Classifier (DTC) as weak learner. \\
We perform experiments on three aspects: \\
1. Weak learner's performance effect. \\
2. Number of weak learner's exploration. \\
\end{abstract}

% Each section begins with a \section{title} command
\section{Introduction}
	% \PARstart{}{} creates a tall first letter for this first paragraph
\PARstart{A}{daboost} short for "Adaptive Boosting",  is an ensemble method that combine a bunch of weak learners into a strong learner which outperforms all the weak learners. It's wildly used in academic and industry community, which make it very important for us to fully explore it. We will explore the the effect of weak learner's performance and the performance with different number of weak learner.

% Main Part
\section{Methods and Theory}
In this part, we define Adaboost algorithm.
\subsection{Adaboost}
Given dataset $ D=\{x_{i}, y_{i}\}_{i=1}^{m}$, where $x_{i} \in \mathbf{R}^{n}$. $y_{i} \in \{-1, 1\}$ and weak Learner $L(x) \in \{-1, 1\} $, Adaboost algorithm can be found in Algorithm~\ref{alg:adaboost}. \par

\begin{algorithm}
	\label{alg:adaboost}
	\caption{Adaboost}
	\KwIn{Training Dataset $ D=\{x_{i}, y_{i}\}_{i=1}^{m}$, num weak learner $N$, weak learner $L$}
	\KwOut{Final learner $H(x)$}
	$w_1(i) \gets \frac{1}{m}$ where i = 1, 2, ..., m \\
	\For{$idx = 1, 2, ..., N$}
	{
		$h_{idx}(x) = L(D, w_{idx})$ \\
		$\epsilon_{idx} = \Sigma_{i=1}^m w_{idx}(i) \mathbb{1}(h_{idx}(x_i) \neq y_i) $ \\
		$\alpha_{idx} = \frac{1}{2} \log{\frac{1 - \epsilon_{idx}}{\epsilon_{idx}}}$ \\
		$w_{idx+1}(i) = w_{idx}(i) * exp(-1 * \alpha_{idx} * y_i * h_{idx}(x_i))$ where i = 1, 2, .., m \\
		$w_{idx+1} = \frac{w_{idx+1}}{\Sigma_{i=1}^m w_{idx}(i)}  $ \\
	}
	return $H(x) = \Sigma_{idx=1}^N \alpha_{idx} h_{idx}(x)$
\end{algorithm} \par

We can see that, in each iteration, the examples that correctly classified by previous weak learner, will get a smaller weight, so that current learner can pay more attention to the misclassified examples.

\section{Experiments}
\subsection{Dataset}
We conduct all the experiments on self constructed face/non-face dataset, which has 500 256x256 face images and 500 32x32 non-face images. All images have 3 channels. We preprocess all the images into 24x24 gray scale images. We then convert them by using NPD features. Finally, each example's form is $(x, y)$, where $x \in \mathbf{R}^{165600}$ and $y \in \{-1, +1\}$, where $+1$ means face and $-1$ means non-face. We randomly split the dataset into three part: 700 training examples, 100 validation examples and 200 test examples. $X_{train},~X_{val}~and~X_{test} $ are divided by 256 for scale purporse.

\subsection{Implementation}
We implement the Adaboost using python and mainly rely on the numpy package.

\subsection{Weak learner's performance's effect.}
In this section, we explore the influence of weak learner's performance in Adaboost. \par 
We first explore the performance of DTC under different depths setting. From Fig~\ref{fig:d_weak_acc} we can see that, as the depth of DTC goes depper, its performance on training set become better but in validation set, the performance of it first gets better and then goes down which indicating the overfitting problem. The best performance in validation set is $89$ occurs when $depth = 4$. \par
Then we explore Adaboost's performance with different weaker learner.  We set number of weak learner to five due to computation limitation. We can see from Fig~\ref{fig:d_acc} that, both Adaboost and DTC suffer from overfitting prolem. But Adaboost outperform DTC under all setting except in validation set when depth equals to 5. We can see that, as the performance of weak learner goes up, the performance of Adaboost goes down, maybe due to the overfitting problem. The intuition is that, if a weak learner's performance is good enough in training set, then Adaboost with more weak learner is prone to overfit the dataset.  \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{d_weak_acc}
		% Create a subtitle for the figure.
		\caption{DTC accuracy under different $depth$.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:d_weak_acc}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{d_acc}
		% Create a subtitle for the figure.
		\caption{Adaboost and DTC accuracy under different $depth$. Note that we set number of weak learner to 5 in Adaboost.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:d_acc}
	\end{center}
\end{figure} \par

\begin{table}[!hbt]
	% Center the table
	\begin{center}
		% Title of the table
		\caption{DTC and Adaboost accuracy on validation set. Best result are in bold.}
		\label{tab:d_dtc_ad}
		% Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
		\begin{tabular}{|c|c|c|c|c|c|}
			\hline
			$depth$  & 1 & 2 & 3 & 4 & 5 \\
			\hline
			$DTC$   & 0.82 & 0.88 & 0.87 & \textbf{0.89} & 0.87   \\
			\hline
			$Adaboost$  & \textbf{0.96} & 0.91 & 0.94 & 0.93 & 0.85 \\
			\hline
		\end{tabular}
	\end{center}
\end{table} \par

\subsection{Performance}
Now we report the final performance of Adaboost under the best settings e.g. DTC'depth = 3 and number of weak learner = 5. The final result is shown in Fig~\ref{tab:final_performance}.

\begin{table}[!hbt]
	% Center the table
	\begin{center}
		% Title of the table
		\caption{Adaboost's best accuracy. }
		\label{tab:final_performance}
		% Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
		\begin{tabular}{|c|c|c|}
			\hline
			$train$  & $validation$ & $test$  \\
			\hline
			1.00   & 0.98 & 0.97  \\
			\hline
		\end{tabular}
	\end{center}
\end{table} \par

\subsection{Number of weak learner's exploration}
In this section, we conduct experiment on number of weak learner's exploration. \par 
We first set the depth to 1 due to computation limitation. Under this setting, More weak Learner benefits the Adaboost, as shown in Fig~\ref{fig:n_ad_acc}.\par
Then we evalutate the performance of Adaboost under settings that DTC' depth ranges from 1 to 5 and number of weak learner ranges from 1 to 5. As illustrated in Fig~\ref{fig:dn_ad_tr_acc}, in training set, Adaboost benefits from both the depth and the number of weak learners. Note that, as the depth of weak learner goes up, the performance gained by adaboost goes down. As shown in Fig~\ref{fig:dn_ad_val_acc}, when DTC's depth equals to 3, as the number of weak learner goes up, Adaboost has better performance. When DTC's depth equals to 5, Adaboost's performance is the worst. \par
Now, we cann draw the following conclusions: \\
1. Weaker learner gets larger performance gain in Adaboost. \\
2. Very Strong weak learner tends to have overfitting problem in Adaboost. \\
3. More weak learner boost the performance in Adaboost when weak learner is not very strong. \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{n_ad_acc}
		% Create a subtitle for the figure.
		\caption{Adaboost accuracy with different number of DTC ($depth=1$).}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:n_ad_acc}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{dn_ad_tr_acc}
		% Create a subtitle for the figure.
		\caption{Adaboost accuracy with different number of DTC ($depth = 1, 2, 3, 4, 5$) in training set.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:dn_ad_tr_acc}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{dn_ad_val_acc}
		% Create a subtitle for the figure.
		\caption{Adaboost accuracy with different number of DTC ($depth = 1, 2, 3, 4, 5$) in validation set.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:dn_ad_val_acc}
	\end{center}
\end{figure} \par

\section{Conclusion}
In this report, we explore Adaboost algorithm. Specifically, we explore the weak learner's performance and number of weak learner in Adaboost. We draw the conclusions that weaker learner can get more performance gain in Adaboost and stronger learner tends to have overfitting problem in Adaboost. And finally we report the performance on test set under the best validation setting. \par

>>>>

% Define document title and author
	\title{Recommender System Based on Matrix Factorization}
	\maketitle

% Write abstract here
\begin{abstract}
In this report, we solve the Recommendation System problem based on Matrix Factorization (MF). \\
We perform experiments on three aspects: \\
1. Hidden dimension's impact. \\
2. Loss function's impact. \\
3. Regularizer's impact. \\
\end{abstract}

% Each section begins with a \section{title} command
\section{Introduction}
	% \PARstart{}{} creates a tall first letter for this first paragraph
\PARstart{M}{atrix} Factorization is a class of collaborative filtering algorithms used in recommender systems. Matrix factorization algorithms work by decomposing the user-item interaction matrix into the product of two lower dimensionality rectangular matrices Usually, the user-item interaction matrix is very sparse e.g. a lot of zeros or mising values in it. The adavantage of MF is that it transforms a sparse matrix into two lower dimensionality matrices which are regarded as User Embeddings (UE) and Item Embeddings (IE) respectively. MF is wildly used in academic and industry community, which make it very important for us to fully explore it. We will explore it under different settings.

% Main Part
\section{Methods and Theory}
In this part, we define Matrix Factorization algorithm.
\subsection{Matrix Factorization}
Given a rating matrix $\mathbf{R} \in \mathbb{R}^{m \times n} $, with sparse ratings from m users to n items. MF factorizes $\mathbf{R}$ into $P$ and $Q$, where $P \in \mathbb{R}^{m \times h }$, $Q \in \mathbb{R}^{n \times h}$ and $h$ is the hidden dimension which is a hyper-parameter. In this way, rating matrix can be expressed as $\mathbf{R} = P Q^{T}$.  Usually, $h \ll min(m, n)$. The Space complexity is $O(m*n)$ for $\mathbf{R}$ and $O((m+n)*h)$ for $P$ and $Q$, where $O((m+n)*h) \ll O(m*n)$. So MF can be used for space saving strategy in some scenarios. Beyond that, the factorized $P$ and $Q$ can be used for unseen user-item pair prediction, which is the core of MF.\par

\section{Experiments}
\subsection{Dataset}
We conduct all the experiments on MovieLens-100k dataset, which contains 10,000 comments from 943 users out of 1682 movies. Each comment's form is $(user_{id}, item_{id}, rating)$, where $ rating \in \{1, 2, 3, 4, 5\}$. We randomly split the dataset into three part: 70,000 training examples, 10,000 validation examples and 20,000 test examples. \par

\subsection{Implementation}
We implement the Matrix Factorization using python and mainly rely on the numpy package. The initializer for $P$ and $Q$ is Xavier uniform initializer, which initial matrix $M$, but bounded to interval $[-a*b, a*b]$, where $n$ is $M \in \mathbb{R}^{a\times b}$. We use Adam optimizer with leraning rate 0.001, set batch size to 256 and epoch to 200. We use $MAE$ as evalutation metric. If not sepecifficly stated, we use $h = 10$, $MSE$ as loss function and $l_2~Regularizer$ as regularizer and penalty cofficient 0.1 in regularizer.

\subsection{Hidden dimension's impact}
In this section, we explore the influence of hidden dimension in MF. We validate the hidden dimension in MF in a range, e.g. 1, 5, 10, 15, 20, 25, 30. \par 
From Fig~\ref{fig:h_mae} we can see that, bigger $h$ tends to have overfitting problem. As presented in Table~\ref{tab:h_mae}, $d = 10$ gets the best performance in validation set. Surprisingly, $h=1$ still get $MAE$ about $0.7522$ in validation set, which is not very far from the best performance e.g. $0.7259$, but $h=1$ can be very time-efficiency and space-saving.  \par
From Fig~\ref{fig:h_tr_losses} and Fig~\ref{fig:h_val_losses} we can see that, bigger $h$ converges faster. \par
\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{h_mae}
		% Create a subtitle for the figure.
		\caption{MF $MAE$ under different $h$.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:h_mae}
	\end{center}
\end{figure} \par

\begin{table}[!hbt]
	% Center the table
	\begin{center}
		% Title of the table
		\caption{MTC $MAE$ under different $h$. Best result are in bold.}
		\label{tab:h_mae}
		% Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
		\begin{tabular}{|c|c|c|c|c|c|c|c|}
			\hline
			$h$  & 1 & 5 & 10 & 15 & 20 & 25 & 30 \\
			\hline
			$MAE_{train}$ & 0.7211 & 0.6216 & 0.5758 & 0.5478 & 0.5302 & 0.5191 & \textbf{0.5124}  \\
			\hline
			$MAE_{val}$  & 0.7522 & 0.7276 & \textbf{0.7259} & 0.7262 & 0.7265 &  0.7321 & 0.7349  \\
			\hline
		\end{tabular}
	\end{center}
\end{table} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{h_tr_losses}
		% Create a subtitle for the figure.
		\caption{MF $MSE_{train}$ under different $d$ and $epoch$. }
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:h_tr_losses}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{h_val_losses}
		% Create a subtitle for the figure.
		\caption{MF $MSE_{val}$ under different $d$ and $epoch$. }
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:h_val_losses}
	\end{center}
\end{figure} \par

\subsection{Loss function's impact}
In this section, we explore the impact of loss function in MF. In particular, we evaluate three loss functions: \\ \\
1. $MAE(y, \hat{y}) = \frac{1}{m}\Sigma_{i=1}^{m}|y_i - \hat{y_i}|$ \\ \\
2. $MSE(y, \hat{y}) = \frac{1}{2m}\Sigma_{i=1}^{m}(y_i - \hat{y_i})^2$ \\ \\
3. $HB(y, \hat{y}) = \frac{1}{m}\Sigma_{i=1}^{m}L_{\delta }(y_i, \hat{y_i}) $ \\ \\
where, $ L_{\delta }(y_i,\hat{y_i})={\begin{cases}{\frac  {1}{2}}(y_i-\hat{y_i})^{2}&{\textrm  {for}}~|y_i-\hat{y_i}|\leq \delta ,\\\delta \,|y_i-\hat{y_i}|-{\frac  {1}{2}}\delta ^{2}&{\textrm  {otherwise.}}\end{cases}}$ \\ \\
We set $\delta = 1$. \par
Three loss functions are mean absolute error, mean squared error and huber loss respectively. The plot of $MAE(x)$, $MSE(s)$ and $HB(x)$ can be found in Fig~\ref{fig:loss_fns}. Huber Loss can been viewed as a combination of $MSE$ and $MAE$. \par
From Fig~\ref{fig:loss_val_mae}, we can see that, HB gets the best performance on validation set. Intuitively, HB takes MAE's adavantage that robust to outliers and MSE's adavantage that smooth in low error region, so it's prone to perform well. The detailed result can be found in Table~\ref{tab:loss_mae}. \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{loss_fns}
		% Create a subtitle for the figure.
		\caption{$MAE$, $MSE$ and $HB$ plot.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:loss_fns}
	\end{center}
\end{figure} \par

\begin{table}[!hbt]
	% Center the table
	\begin{center}
		% Title of the table
		\caption{MF's performance under different loss function settings. Best result are in bold.}
		\label{tab:loss_mae}
		% Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
		\begin{tabular}{|c|c|c|c|}
			\hline
			$loss_{fn}$  & $MAE$ & $MSE$ & $HB$  \\
			\hline
			$MAE_{val}$ & 0.7293 & 0.7244 & \textbf{0.7217} \\
			\hline
		\end{tabular}
	\end{center}
\end{table} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{loss_val_mae}
		% Create a subtitle for the figure.
		\caption{MF $MAE_{val}$ under different loss functions.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:loss_val_mae}
	\end{center}
\end{figure} \par

\subsection{Regularizer's impact}
In this section, we explore the impact of regularizer in MF. In particular, we evaluate three regularizers: \\ \\
1. $L1(w) = \lambda * \Sigma_{i} |w_i| $ \\ \\
2. $L2(w) = \lambda * \Sigma_{i} w_i^2 $ \\ \\
3. $L1\_L2(w) = \frac{L1(w) + L2(w)}{2}  $ \\ \\
We set $\lambda = 0.01$ in all experiments. \par
$L1\_L2$ regularizer can be seen as a combination of $L1$ and $L2$ regularizer. The plot of three regularizers can be found in Fig~\ref{fig:regs}.\par
From Fig~\ref{fig:reg_val_mae}, we can see that, $L2$ gets the best performance on validation set. $L1\_L2$ has sever overffiting problem than $L1$ and $L2$ after many epoches. However, $L1\_L2$ gets better performance than $L1$.  \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{regs}
		% Create a subtitle for the figure.
		\caption{$L1$, $L2$ and $L1\_L2$ plot.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:regs}
	\end{center}
\end{figure} \par

\begin{table}[!hbt]
	% Center the table
	\begin{center}
		% Title of the table
		\caption{MF's performance under different regularizer settings. Best result are in bold.}
		\label{tab:reg_mae}
		% Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
		\begin{tabular}{|c|c|c|c|}
			\hline
			$Regularizer$  & $L1$ & $L2$ & $L1\_L2$  \\
			\hline
			$MAE_{val}$ & 0.7537 & \textbf{0.7243} & 0.7358 \\
			\hline
		\end{tabular}
	\end{center}
\end{table} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{reg_val_mae}
		% Create a subtitle for the figure.
		\caption{MF $MAE_{val}$ under different regularizer settings.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:reg_val_mae}
	\end{center}
\end{figure} \par

\subsection{Final Performance}
In this section, we report the final performance under best settings, e.g. $h = 10$, $loss_{fn} = HB$ and $regularizer = L2$. Best result can be found in Table~\ref{tab:final_performance}. \par
\begin{table}[!hbt]
	% Center the table
	\begin{center}
		% Title of the table
		\caption{MF's best performance under best settings.}
		\label{tab:final_performance}
		% Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
		\begin{tabular}{|c|c|c|}
			\hline
			$MAE_{train}$  & $MAE_{val}$ & $MAE_{test}$ \\
			\hline
		      0.6342 & 0.7246 & 0.7294\\
			\hline
		\end{tabular}
	\end{center}
\end{table} \par

\section{Conclusion}
In this report, we explore Matrix Factorization algorithm. Specifically, we explore the hidden dimension $h$, loss function and regularizer in MF. We oberseve that smaller $h$ causes under-fitting problem while bigger $h$ causes over-fitting problem. Under the settings that using Huber Loss as loss function and L2 as regularizer, the best perfomance can be obtained.   \par
